### 2017-12-19 paper/paper.tex
very rough intro/related work

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}

%
\begin{document}
%
\mainmatter              % start of the contributions
%
\title{An Executable Formal Semantics For A Functional Actor Language}
%
\author{Xiaohong Chen, Lucas Pe\~{n}a}
%
\institute{University of Illinois at Urbana-Champaign, \\
\email{\{xc3, lpena7\}@illinios.edu}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
A simple reference actor language has been proposed in~\cite{}, serving as a
sound mathematical calculus foundation for various actor languages used in
distributed applications that involves concurrency. Even though the authors
of~\cite{} provided in the paper an operational semantics of the actor language,
an executable implementation was not given. In this project, we will give the
reference actor language an implementation that by construction conforms to its
formal operational semantics using the K framework~\cite{}, a rule-based
semantic framework. We will show that our executable formal semantics are
capable of executing some meaningful actor systems examples. We hope our
executable formal semantics can be served as a starting point for future work in
developing automatic equivalence provers using the K framework.
\end{abstract}
%
\section{Introduction}
The K framework~\cite{} is a framework for defining executable operational
semantics of a programming language. K is based on rewriting logic, organizing
the state of a program into cells and rewriting the particular parts of the
configuration. K also has access the entire computation at any particular
time. These features and more allow for difficult language constructs to be
designed with ease in K. On top of that, K provides backend tooling for
additional features based on the semantics of a programming language, such as a
semantic debugger and a full formal verifier. See Figure~\ref{fig:k-tools} for a
complete view of the tools one gets from K based off an operational semantics.

\subsection{Related Work}

\begin{figure}
          \centering
          \includegraphics[scale=0.13]{../k-overview.png}
	  \caption{Backend tools K provides from just an operational
            semantics. Some tools such as a semantics-based compiler and
            test-case generation are in progress.}
          \label{fig:k-tools}
\end{figure}
To demonstrate the power of K, languages as complex as C, Java, and JavaScript
have been fully defined using the K framework~\cite{ , , }. On top of that, in
all of these languages and more, nontrivial programs have been formally verified
using K's prover, such as proofs involving AVL trees and red-black trees. We
would like to add an actor-based language to the growing list of languages
defined in K, so we can take advantage of K's formal verification engine and
more.

Semantics of actor-based languages have been defined on paper, for example
in~\cite{}. Additionally, there are many implementations of actor-based systems,
including Erlang~\cite{}, Akka~\cite{}, Scheme~\cite{}, and much more. However,
there is currently no formally defined \textit{executable} operational semantics
of an actor-based language. In K, there was previous work done to define the
semantics of Orc, a concurrent programming language~\cite{}, though that work is
incomplete.

In this paper, we aim to define the semantics of the actor language proposed
in~\cite{}.  We choose this language as the semantics are clearly defined in the
paper, and it is one of the more simple actor frameworks. Defining this language
in K will give assurance to the correctness of the language - as K is executable
this can be tested much easier than defining a language on paper. It will also
provide more assurance on K's language-defining ability, as well as open up to
formal analysis using the tools seen in Figure~\ref{fig:k-tools}.

\section{The Actor Language in K}
\subsection{Configuration}
\subsection{Syntax and Evaluation Contexts}
\subsection{Semantics}
\subsection{In-depth Example}

\section{Conclusion and Future Work}

%
% ---- Bibliography ----
%
\bibliographystyle{ieeetr}

\bibliography{references}

\end{document}